### 2016-02-02 paper/bennett_leonard.tex

\documentclass{article}

\usepackage{graphicx}
\usepackage{float}
\usepackage{subcaption}
\usepackage{microtype}

\begin{document}

\title{On the Statistics of Macrospicules}
\author{S.M.Bennett$^1$
\author{A. Leonard}
and
\and R. Erd{\'e}lyi$^{1,2}$}
\affil{$^1$ Solar Physics and Space Plasma Research Centre (SP2RC), University of Sheffield, Hicks Building, Hounsfield Rd, S3 7RH, UK}
\affil{$^2$ Institute of Maths, Physics and Computer Science (IMPACS), Aberystwyth University, Aberystwyth, Ceredigion, SY23 3FL}
\affil{$^3$ Debrecen Heliophysical Observatory (DHO), Research Centre for Astronomy and Earth Sciences, Hungarian Academy of Sciences, 4010 Debrecen, P.O. Box 30, Hungary}