### 2016-04-04 ms_temp_paper.tex

% DOCUMENT PREAMBLE
\documentclass{emulateapj}
% \documentclass[referee]{apj}

\usepackage{graphicx}
\usepackage{float}

\usepackage{subcaption}

\usepackage{apjfonts}
\usepackage{microtype}

\begin{document}

\title{Solar jet observed at the limb in SDO/SST}
\author{S.M.Bennett$^1$
and
\author{A.J.Leonard$^1$}
\affil{$^1$ Solar Physics and Space Plasma Research Centre (SP2RC), University of Sheffield, Hicks Building, Hounsfield Rd, S3 7RH, UK}

\begin{abstract}

\end{abstract}

\maketitle

\section{Introduction}

\section{Observation}

\section{Results}

\section{Discussion}

\section{Conclusion}

\end{document}